---
title: "COMP1013 Analytics Programming Project"
author: "Phạm Minh Khôi – 22145599"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output: 
  pdf_document:
    number_sections: true
fontsize: 12pt
geometry: margin=1in
---
## Q1 — Data Cleaning and Preprocessing

In this step, all three datasets were inspected for non-standard missing value encodings.  
The datasets used `"?"` as a missing marker in categorical fields. These must be standardised to real `NA` values so that statistical functions handle them correctly. Key categorical variables are converted into factors because they represent discrete categories rather than continuous values. Horsepower has missing values, therefore median imputation was used, which is robust to outliers and preserves distribution central tendency.

```{r}
# Load CSV data

engines     <- read_csv("Engine.csv")
automobile  <- read_csv("Automobile.csv")
maintenance <- read_csv("Maintenance.csv")

head(engines)
head(automobile)
head(maintenance)

```

```{r}
# count rows containing '?' before replacement
rows_with_q <- function(df) {
  char_df <- dplyr::select(df, where(is.character))
  if (ncol(char_df) == 0) return(0L)
  has_q <- apply(char_df, 1, function(row) any(row == "?", na.rm = TRUE))
  sum(has_q)
}
affected_engines_rows     <- rows_with_q(engines)
affected_automobile_rows  <- rows_with_q(automobile)
affected_maintenance_rows <- rows_with_q(maintenance)

# replace '?' with real NA
replace_q_to_na <- function(df){
  df %>% mutate(across(where(is.character), ~na_if(., "?")))
}
engines_na     <- replace_q_to_na(engines)
automobile_na  <- replace_q_to_na(automobile)
maintenance_na <- replace_q_to_na(maintenance)

# convert categoricals to factors
automobile_na$BodyStyles  <- factor(automobile_na$BodyStyles)
engines_na$FuelTypes      <- factor(engines_na$FuelTypes)
maintenance_na$ErrorCodes <- factor(maintenance_na$ErrorCodes)

# impute Horsepower with median
engines_before_impute <- engines_na
engines_na$Horsepower <- as.numeric(engines_na$Horsepower)
hp_median <- median(engines_na$Horsepower, na.rm = TRUE)
engines_na$Horsepower[is.na(engines_na$Horsepower)] <- hp_median
```

```{r}
# impact summary table (missing replaced + verification)
impact_df <- data.frame(
  Table = c("Engines","Automobile","Maintenance"),
  Rows_with_Q_before = c(affected_engines_rows,
                         affected_automobile_rows,
                         affected_maintenance_rows)
)
impact_df$Total_Q_after <- c(
  rows_with_q(engines_na),
  rows_with_q(automobile_na),
  rows_with_q(maintenance_na)
)
impact_df
```

```{r}
# Horsepower distribution after imputation
ggplot(engines_na, aes(Horsepower)) +
  geom_histogram(bins=30) +
  labs(title="Horsepower Distribution After Median Imputation",
       x="Horsepower", y="Count") +
  theme_minimal()
```
